### 2021-06-10 results.tex
update on the response letter to address some observation from R3
presenting a better motivation for the pieces of code

\begin{finding}
  Our results suggest that one might benefit from using  an integrated
  environment that combines Monkey, DroidMate, and Humanoid to
  mine Android sandboxes. Introducing the DroidMate 
  tool does not improve the results.
\end{finding}

@@

Altogether, ignoring the \joke tool, our study reveals that from $58.33$\% (Humanoid)
to $76.04$\% (DroidBot) of the malicious apps investigated in our study can be
detected using the sandboxes generated after running the test case tools with the support of the
DroidFax static analysis algorithms. Besides that, in the first execution (WS), none of the
resulting sandboxes could detect 18 malwares in our dataset ($18.75$\%). According to
the Euphony tool~\cite{hurier2017euphony}, $12$ of these $18$ malwares are \emph{adwares}, $3$ are \emph{trojans}, $2$ are
PUPs (\emph{Potentially Unwanted Program}), and one is an \emph{exploit}. In what follows
we present more details of representative malwares considered
in our dataset.

First, consider the malicious version of the app \texttt{com.andoop.flyracing}---which only the DroidBot and
Humanoid sandboxes were able to detect. 
In this particular case, the malicious version changed the Android Manifest file,
adding permissions to receive and send SMS messages
(Listing~\ref{lst:androidManifest}). With these permissions, malicious apps may get money
fraudulently by sending messages without user confirmation, or may monitor messages or delete
them without showing them to user.

Moreover, after decompiling this malware, we also observed that the malicious version of the
\texttt{MainService} class introduced a
behavior that collects sensitive information (the International Mobile
Equipment Identity, IMEI) and sends it using an SMS message
(Listing~\ref{lst:mainService}).

@@

The malicious version of the app \texttt{com.happymaau.MathRef} also changes
the manifest file to require additional permissions as well as change
the behavior of the app (with malicious code). All sandboxes were able to
detect this malware.
In this case, the malicious version of the app changed the Android Manifest file,
requiring permissions to access the network and WiFi states (Listing~\ref{lst:androidManifest2}).
These changes allow an app
to view the status of all networks and make changes to configured WiFi-fi networks.

@@

We also observed that the malicious version introduces a method \texttt{a},
that actually collects network and WiFi information, like Mac address and the network state
(see Listing~\ref{lst:d}). This information is then shared using an
HTTP request.

@@

All computed sandboxes also detected the malicious version of the app \texttt{ru.qixi.android.smartrabbits}.
This particular malware also changes the Android Manifest file, requesting permission to access location service (Listing~\ref{lst:androidManifest3}).
This permission allows access location features, such as the Global Positioning System (GPS) on the phone, if it is enabled.
Malicious applications can use these features to determine where the phone owner is, which is a
general privacy threat.

@@

In addition, the malicious app clandestinely monitors the geographic location of the user and sink
these information to a web server. Listing~\ref{lst:c} shows how
the method \texttt{c}, from \texttt{q} class, collects these sensitive information.

@@

Finally, the \texttt{com.andoop.flyracing} case is among the apps that none sandbox detected the
malicious version. Indeed, the malicious version only changes the Android Manifest file,
modifying the meta-data \texttt{ADMOB\_PUBLISHER\_ID}. Nonetheless, the AdMob is a monetizing
service provided by Google, and changing the AdMob \emph{publisher identifier} account redirects
the advertisement's revenue to another destination. Therefore, it might exist specific
modifications to the Android manifest file that might complement the mining sandbox
approach into the task for detecting malwares.

@@

\subsection{Results of the second study: Using taint analysis to mine Android malwares}\label{sec:res-ss}

>>>>

\begin{finding}
  Our results suggest that one might benefit from using  an integrated
  environment that combines Monkey, DroidMate, and Humanoid to
  mine Android sandboxes. Contrasting with the \blls, introducing the DroidMate 
  tool does not improve the overall performance for detecting malwares using
  a mining sandbox approach.
\end{finding}

@@

Altogether, ignoring the \joke tool, our study reveals that from $58.33$\% (Humanoid)
to $76.04$\% (DroidBot) of the malicious apps investigated in our study can be
detected using the sandboxes generated after running the test case tools with the support of the
DroidFax static analysis algorithms. Besides that, in the first execution (WS), none of the
resulting sandboxes could detect 18 malwares in our dataset ($18.75$\%). According to
the Euphony tool~\cite{hurier2017euphony}, $12$ of these $18$ malwares are \emph{adwares}, $3$ are \emph{trojans}, $2$ are
PUPs (\emph{Potentially Unwanted Program}), and one is an \emph{exploit}.
At this point, an additional question arises: what are the characteristics
of the malwares that have (not) been identified using the mining sandbox approach?
To explore this question, we take advantage of the
\texttt{dex2jar} tool to reverse-engineer all 96 malwares considered in our analysis and
computed the \emph{diffs} of the benign/malicious versions of the APPs.
The results of this activity are available in our replication
package.\footnote{https://github.com/droidxp/paper-replication-package/blob/master/diff/} 
In what follows we dissect a few examples of malwares
that at least one dynamic analysis tool was able to
identify using the mining sandbox approach. After that,
we present the characteristics of a malware that none of the tools
was able to detect. Our goal here is to provide
a lower-level intuition about the classes of malware the
mining sandbox approach is able to detect. A reader that
is not interested in these details could skip to
Section~\ref{sec:res-ss}.

To start with, consider the malicious version of the app \texttt{com.andoop.flyracing}---which both DroidBot and
Humanoid sandboxes could detect in our analysis. 
In this particular case, the malicious version changes the Android Manifest file,
adding permissions to receive and send SMS messages
(Listing~\ref{lst:androidManifest}). With these permissions, malicious apps may get money
fraudulently by sending messages without user confirmation, for instance.

After decompiling this malware, we also observed that the malicious version of the
\texttt{MainService} class introduces a
behavior that collects sensitive information (the International Mobile
Equipment Identity, IMEI) and sends it using an SMS message
(Listing~\ref{lst:mainService}).

@@

The malicious version of the app \texttt{com.happymaau.MathRef} also changes
the manifest file to require additional permissions as well as change
the behavior of the app (with malicious code). All sandboxes were able to
detect this malware.
In this case, the malicious version of the app changes the Android Manifest file,
requiring permissions to access the network and WiFi states (Listing~\ref{lst:androidManifest2}).
These changes allow an app
to view the status of all networks and make changes to configured WiFi networks.

@@

The malicious version also introduces a method \texttt{a},
that actually collects network and WiFi information, like Mac address and the network state
(see Listing~\ref{lst:d}). This information is then shared using an
HTTP request.

@@

All resulting sandboxes also detected the malicious version of the app \texttt{ru.qixi.android.smartrabbits}.
This particular malware also changes the Android Manifest file,
requesting permission to access the location service (Listing~\ref{lst:androidManifest3}).
This permission allows access to location features, such as the Global Positioning System (GPS) on the phone, if it is enabled. Malicious applications can use these features to determine where the phone owner is, which is a
classic and well-documented privacy threat.

@@

In addition, the malicious app clandestinely monitors the geographic location of the user and sink
this information to a web server. Listing~\ref{lst:c} shows how
the method \texttt{c}, from the class named \texttt{q}, collects this sensitive information.

@@

This pattern of changing the Android Manifest file and including
new method calls characterizes the classes of malwares for which
the mining sandbox approach excels. 
In a different vein, the malicious version of the app
\texttt{com.andoop.flyracing} is among the apps that none of the sandboxes could
detect. Indeed, the malicious version only changes the Android Manifest file,
modifying the meta-data \texttt{ADMOB\_PUBLISHER\_ID}. The AdMob is a monetizing
service provided by Google, and changing the AdMob \emph{publisher identifier} account redirects
the advertisement's revenue to another destination. Based on
this observation, we envision integrating a different approach that reasons
about modifications to the Android manifest file and that might complement the mining sandbox
approach into the task for detecting malwares; since the mining
sandbox approach is not able to detect malicious packages that
do not introduce new method calls for sensitive APIs.

@@

\subsection{Results of the second study: Use of Taint Analysis for Malware Identification}\label{sec:res-ss}
